Comprehensive quality - aware Automated Semantic Web Service Composition

### 2016-11-12 paper.tex
update

\section{Comprehensive quality-aware Semantic Automated Web Service Composition}\label{qswsc_approach}
\begin{figure}[h]
\centerline{
\fbox{
\includegraphics[width=12cm]{overview.pdf}}}
 \caption{Overview of the proposed approach.}
 \label{overview}
\end{figure}
In this paper, we propose a comprehensive quality model for automated semantic service composition, and optimise both the quality of semantic matchmaking and QoS. PSO has shown its efficiency in solving combinatorial optimisation problems. Therefore, we will employ a PSO-based approach, which is considered to be simple and efficient without penalising or repairing that often required by GP \cite{da2014graph}. Fig. \ref{overview} shows the overview of our approach with five steps. Step 1: The composite process is triggered by a composite goal defined in Subsection \ref{problemDes}, which describes customers requirements both functional and non-functional. Step 2: This composition goal are used to discover all relevant web services, which lead to a shrunken service repository that is subsequently used by PSO as a searching space. Step 3: A weighted graph representation is randomly built up from the an initial service queue that mapped to the particle's location, interleaving with semantic matchmaking process. In the weighted graph, graph edges are assigned with semantic matchmaking quality as weights. Step 4: The fitness value of the weighted graph is evaluated to update the position of particle under PSO algorithm in Sect. \ref{pso_algorithm}, where the position is mapped to the index of service queue. later on, the updated service queue is used to decode a new weighted graph as the composition solution. Step 4. Lastly, the best position found in the searching space is selected and decode into the final optimised solution. This PSO-based approach is different from \cite{da2016particle}, as we use weighted graphs as solution presentation.

@@

The overall algorithm investigated here is made up of a PSO-based web service composition algorithm \ref{novelSteps} and a decoding algorithm \ref{graph_building}, to generate a composition weighted graph. In algorithm \ref{novelSteps}, the idea is to translate the particle location produced by PSO into a service queue as an indirect representation of weighted graph, such that finding the best fitness of the weighted graph is to discover the optimised location of the particle in the search space. In PSO, the dimension of each particle is equals to the number of relevant web services. The index of each services is mapped to a separate location component in a particle. Services in a queue follow the ascending order, from which we decode a weighted graph using Algorithm \ref{graph_building}. We select and connect  service vertex to the start vertex from the service queue if the web service can satisfy the output of start vertex. Meanwhile, an Output Set is initialised and kept being updated using outputs from the newly added service vertex in the weighted graph, later on, more services vertices are connected with edges if services' inputs could be satisfied by any output in the Output Set. Finally, the end vertex is connected to the graph if the updated Output Set contains all end inputs. In addition, dangling service vertex and edges will be removed.

@@

We run the experiment on computing grid comprising of almost 170 NetBSD (Unix operating system) workstations operated by the Sun Grid Engine. Experimentation is done using the approach explained in Section \ref{qswsc_approach}, and evaluated in the comprehensive quality model introduced in Subsection \ref{qualityModel}. The parameters were chosen based on general settings from \cite{shi2001particle} for our PSO-based approach, 30 particles evolved in 100 generations in the searching space, in which problem dimension's size equals to the number of relevant services in service repository. We run 30 times independently for each dataset. We configure weight of fitness function to properly balance functional side and nonfunctional side. Therefore, $w_{1}$ and $w_{2}$ are equals to 0.1 and 0.4,  and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125 accordingly. However, the settings of the values does not impact the method. In general, weight settings are adjusted according to users' preferences.

@@

In this section, we analyse the composition solution generated by using our approach comparing with QoS-aware approach. In particularly, we show that our proposed approach can produce solutions that results better matchmaking, which meets user's goal better.

First, we look at mean value of $Q_{mt}$, $Q_{s}$ and $Q_{QoS}$ at optimum at the 100th generation for two approaches, shown in Table \ref{decisionTable}, where QoS-aware approach record $Q_{mt}$, $Q_{s}$ while utilise fitness function $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$ when $\sum_{i=1}^{4} w_i = 1$ without considering $MT$ and $S$, and $Q_(QoS)$ is normalised from 0 to 0.5 to make it comparable to $Q_{QoS}$ in our approach. We observe an interesting pattern from the Table \ref{decisionTable} using statistic analysis: mean value of $Q_{mt}$ and $Q_{s}$ at optimum by our approach is consistently higher than those by QoS-aware approach in Task 1, 3, 4, and 5 with an exception in Task 2 while the combination of mean $Q_{mt}$ and $Q_{s}$ at optimum in Task 2 is still considered to be higher in our approach with 0.1 and 0.4 weights respectively. Meanwhile, $Q_{QoS}$ generated by QoS-aware approach can obtain a slightly higher value than that of our approach. In conclusion, we can perceive that our evaluation could find out better functional quality with a reasonable trade off in QoS.

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table}[]
\footnotesize
\centering
\caption{Mean Quality for comprehensive quality-aware methods and QoS-aware approach}
\label{decisionTable}
\begin{tabular}{l|l|l|l}
\hline
\multicolumn{2}{c|}{WSC09}              & \shortstack{QoS-aware \\ Evaluation} & \shortstack{Comprehensive \\ Quality Evaluation} \\ \hline
\multirow{3}{*}{Task1}  &$Q_{mt}$   &0.189787 $\pm$ 0.039278686  &0.220594 $\pm$ 0.009318 $\uparrow$ \\ \cline{2-4} 
                        &$Q_{s}$    &0.884962 $\pm$ 0.014140   &0.893713 $\pm$ 0.007769  $\uparrow$ \\ \cline{2-4}
                        &$Q_{QoS}$  &0.278730 $\pm$ 0.007786     &0.279548 $\pm$ 0.007627                          \\ \hline
\multirow{3}{*}{Task2}  &$Q_{mt}$   &0.001795 $\pm$ 0.000719 $\uparrow$    &0.001135 $\pm$ 0.000633  \\ \cline{2-4} 
                        &$Q_{s}$    &0.906971 $\pm$ 0.005855     &0.927965 $\pm$ 0.007306 $\uparrow$   \\ \cline{2-4}
                        &$Q_{QoS}$  &0.239979 $\pm$ 0.000578 $\uparrow$     &0.237845 $\pm$ 0.001439 \\ \hline
\multirow{3}{*}{Task3}  &$Q_{mt}$   &0.158526 $\pm$ 0.014028               &  0.245830 $\pm$ 0.007761 $\uparrow$   \\ \cline{2-4} 
                        &$Q_{s}$    &0.949109 $\pm$ 0.002331           &   0.972765 $\pm$ 0.002980 $\uparrow$           \\ \cline{2-4}
                        &$Q_{QoS}$  &0.247002 $\pm$ 0.000661 $\uparrow$             &  0.245631 $\pm$ 0.000431            \\ \hline
\multirow{3}{*}{Task4}  &$Q_{mt}$   &0.000000 $\pm$ 0.000000  &0.000001 $\pm$ 0.000002 $\uparrow$ \\ \cline{2-4} 
                        &$Q_{s}$    &0.880058 $\pm$ 0.006603    &0.930632 $\pm$ 0.009043 $\uparrow$ \\ \cline{2-4}
                        &$Q_{QoS}$  &0.242338 $\pm$ 0.000516 $\uparrow$    &0.236677 $\pm$ 0.002211  \\ \hline
\multirow{3}{*}{Task5}  &$Q_{mt}$   &  0.000042 $\pm$ 0.000030     &  0.000078 $\pm$ 0.000031 $\uparrow$           \\ \cline{2-4} 
                        &$Q_{s}$    &  0.915933 $\pm$ 0.012888      &  0.928189 $\pm$ 0.002486 $\uparrow$                 \\ \cline{2-4}
                        &$Q_{QoS}$  & 0.238189 $\pm$ 0.000240$\uparrow$      &  0.237232 $\pm$  0.000391  \\ \hline                                                   
\end{tabular}
\end{table}

Second, to compare the results generated from two evaluation approaches, we demonstrate an example solution that shows the differences in two composition weighted graphs representations under two different methods. The Fig. \ref{comparison} shows two composition weighted graphs as the solutions to Task 3 with $(1)$ QoS-aware approach and $(2)$ Comprehensive quality-aware method respectively. Two approaches generate exactly the same service workflow structure while some service vertex and edges denoted in red are different. Therefore, we make a comparison of the quality among these different edges ($e_1$ to $e4$) associated service vertex in terms of quality of semantic matchmaking and QoS attributes. We also look at $\Delta Q$ which reveals the amount of variation on quality between two methods, where the positive values means the benefits gained under our approach while the negative values means the trade-off. To demonstrate the benefits of the positive value, we analyse the smallest positive $\Delta Q$ corresponding to $e_4$ and demonstrate how the output and required input are different under two approaches in in Fig. \ref{comparisontaxonomy}. $Ser1640238160$ and $Ser1947554374$ are selected service vertex with output concept-related parameters $Inst795998200$ and $Inst582785907$ corresponding to $Con103314376$ and $Con2037585750$ respectively while $Inst658772240$ are the required input related to $Con2113572083$, in which there exist $Con103314376$ $\sqsubseteq_{2}$ $Con2113572083$ and $Con2037585750$ $\sqsubseteq_{3}$ $Con2113572083$. It is obviously that our approach select the service providing $Inst795998200$ that are closer the users' requirement, which means our methods results in better semantic matchmaking quality.

@@

To analyse the effectiveness of our approach, we will study the convergence rate of the proposed method in this section to understand the convergence rate of five tasks in WSC09. To further study the overall effectiveness of evaluation model and explore impacts of all involved components in the fitness function. We analysis all the performances of these components in the whole evolutionary process in Fig. \ref{exp_fitnessvalue}, in which the five tasks' experiment results are arranged in four groups consisting of average fitness, average matchType quality, average similarity quality and average QoS for generation 0-99 with optimum.

Firstly, the average fitness value with optimum is calculated by the average value of best fitness found in each generation over 30 independent runs. The value plotted in the first column group ranges from 0 to 1 in Fig. \ref{exp_fitnessvalue}, and it is considered to be more optimised if its value is closer to value 1. We can see that there is a significant increase in the fitness values with optimum between generation 0 and generation 15-25, the remaining generation continues to produce a steady but moderate improvement in the fitness value and eventually reach a plateau with no further changes. The same behaviour is observed over the rest tasks. We also witnessed a fast convergence rate of the fitness value.

We also investigate the variation of quality of semantic matchmaking where average matchType quality with optimum and average similarity quality with optimum are studied from generation 0 to 99 in second and third column groups of Fig. \ref{exp_fitnessvalue}. In these groups,  both the average matchType quality with optimum and average similarity quality with optimum are the mean value corresponding to the best fitness value found in each generation over 30 independent runs. The dominant tendency of the marked value representing a similar characteristic compared to the average fitness value group trend in five tasks. However, there are some slight fluctuations over the generations.

Lastly, it is interesting to look at the average QoS per generation for the 30 independent runs when the functional quality heads up. Here average QoS with optimum is also the value corresponding to the fitness value with optimum in each generation, and we display the results in the fourth column group in Fig. \ref{exp_fitnessvalue}. It is obvious that the overall trend of QoS moves upward to a high constant level in five independent tasks. Additionally, we don't see too much trade-off from the QoS.

>>>>

\section{Comprehensive quality-aware Semantic Automated Web Service Composition}\label{qswsc_approach}
\begin{figure}[h]
\centerline{
\fbox{
\includegraphics[width=12cm]{overview.pdf}}}
 \caption{Overview of the proposed approach.}
 \label{overview}
\end{figure}
In this paper, we propose a comprehensive quality model for automated semantic service composition, and optimise both the quality of semantic matchmaking and QoS. PSO has shown its efficiency in solving combinatorial optimisation problems. Therefore, we will employ a PSO-based approach, which is considered to be simple and efficient without penalising or repairing that often required by GP \cite{da2016particle}. Fig. \ref{overview} shows the overview of our approach with five steps. Step 1: The composite process is triggered by a composite goal defined in Subsection \ref{problemDes}, which describes customers requirements both functional and non-functional. Step 2: This composition goal are used to discover all relevant web services, which lead to a shrunken service repository that is subsequently used by PSO as a searching space. Step 3: A weighted graph representation is randomly built up from the an initial service queue that mapped to the particle's location, interleaving with semantic matchmaking process. In the weighted graph, graph edges are assigned with semantic matchmaking quality as weights. Step 4: The fitness value of the weighted graph is evaluated to update the position of particle under PSO algorithm in Sect. \ref{pso_algorithm}, where the position is mapped to the index of service queue. later on, the updated service queue is used to decode a new weighted graph as the composition solution. Step 4. Lastly, the best position found in the searching space is selected and decode into the final optimised solution. This PSO-based approach is different from \cite{da2016particle}, as we use weighted graphs as solution presentation.

@@

The overall algorithm investigated here is made up of a PSO-based web service composition algorithm \ref{novelSteps} and a decoding algorithm \ref{graph_building}, to generate a composition weighted graph, where code lines in blue are different from the algorithm in \cite{da2016particle}. In algorithm \ref{novelSteps}, the idea is to translate the particle location produced by PSO into a service queue as an indirect representation of weighted graph, such that finding the best fitness of the weighted graph is to discover the optimised location of the particle in the search space. In PSO, the dimension of each particle is equals to the number of relevant web services. The index of each services is mapped to a separate location component in a particle. Services in a queue follow the ascending order, from which we decode a weighted graph using Algorithm \ref{graph_building}. We select and connect  service vertex to the start vertex from the service queue if the web service can satisfy the output of start vertex. Meanwhile, an Output Set is initialised and kept being updated using outputs from the newly added service vertex in the weighted graph, later on, more services vertices are connected with edges if services' inputs could be satisfied by any output in the Output Set. Finally, the end vertex is connected to the graph if the updated Output Set contains all end inputs. In addition, dangling service vertex and edges will be removed.

@@

We run the experiment on computing grid comprising of 170 NetBSD (Unix operating system) workstations operated by the Sun Grid Engine. The parameters were chosen based on general settings from \cite{shi2001particle} for our PSO-based approach, In particular, PSO population size is 30 with 100 generations. We run 30 times independently for each dataset. We configure weight of fitness function to properly balance functional side and nonfunctional side. Therefore, $w_{1}$ and $w_{2}$ equal to 0.1 and 0.4,  and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125 accordingly. However, the settings of these weight values does not impact the method. In general, weight settings are adjusted according to users' preferences.

@@

In this section, we analyse the composition solution generated by using our approach comparing with QoS-aware approach. In particularly, we show that our proposed approach can produce solutions that results in better matchmaking, which meets user's goal better.

First, we look at mean value of $Q_{mt}$, $Q_{s}$ and $Q_{QoS}$ at optimum at the 100th generation for two approaches, shown in Table \ref{decisionTable}. The QoS-aware approach record $Q_{mt}$, $Q_{s}$ and utilise fitness function $Fitness = w_1 \hat{A} + w_2 \hat{R} + w_3(1 - \hat{T}) + w_4(1 - \hat{C})$ when $\sum_{i=1}^{4} w_i = 1$ without considering $MT$ and $S$, in which $Q_(QoS)$ is normalised from 0 to 0.5 to make it comparable to $Q_{QoS}$ in our approach. We observe an interesting pattern from the Table \ref{decisionTable} using statistic analysis: mean value of $Q_{mt}$ and $Q_{s}$ at optimum by our approach is consistently higher than those by QoS-aware approach with one exception in Task 2 while the combination of mean $Q_{mt}$ and $Q_{s}$ at optimum in Task 2 is still significantly higher in our approach under statistical anlysis. Meanwhile, $Q_{QoS}$ generated by QoS-aware approach can obtain a slightly higher value than that of our approach. In conclusion, we can perceive that our evaluation could find out better functional quality with a reasonable trade off in QoS.

\begin{table}[]
\footnotesize
\centering
\caption{Mean Quality for comprehensive quality-aware methods and QoS-aware approach}
\label{decisionTable}
\begin{tabular}{l|l|l|l}
\hline
\multicolumn{2}{c|}{WSC09}              & \shortstack{QoS-aware \\ Evaluation} & \shortstack{Comprehensive \\ Quality Evaluation} \\ \hline
\multirow{3}{*}{Task1}  &$Q_{mt}$   &0.189787 $\pm$ 0.039278686  &0.220594 $\pm$ 0.009318 $\uparrow$ \\ \cline{2-4} 
                        &$Q_{s}$    &0.884962 $\pm$ 0.014140   &0.893713 $\pm$ 0.007769  $\uparrow$ \\ \cline{2-4}
                        &$Q_{QoS}$  &0.278730 $\pm$ 0.007786     &0.279548 $\pm$ 0.007627                          \\ \hline
\multirow{3}{*}{Task2}  &$Q_{mt}$   &0.001795 $\pm$ 0.000719 $\uparrow$    &0.001135 $\pm$ 0.000633  \\ \cline{2-4} 
                        &$Q_{s}$    &0.906971 $\pm$ 0.005855     &0.927965 $\pm$ 0.007306 $\uparrow$   \\ \cline{2-4}
                        &$Q_{QoS}$  &0.239979 $\pm$ 0.000578 $\uparrow$     &0.237845 $\pm$ 0.001439 \\ \hline
\multirow{3}{*}{Task3}  &$Q_{mt}$   &0.158526 $\pm$ 0.014028               &  0.245830 $\pm$ 0.007761 $\uparrow$   \\ \cline{2-4} 
                        &$Q_{s}$    &0.949109 $\pm$ 0.002331           &   0.972765 $\pm$ 0.002980 $\uparrow$           \\ \cline{2-4}
                        &$Q_{QoS}$  &0.247002 $\pm$ 0.000661 $\uparrow$             &  0.245631 $\pm$ 0.000431            \\ \hline
\multirow{3}{*}{Task4}  &$Q_{mt}$   &0.000000 $\pm$ 0.000000  &0.000001 $\pm$ 0.000002 $\uparrow$ \\ \cline{2-4} 
                        &$Q_{s}$    &0.880058 $\pm$ 0.006603    &0.930632 $\pm$ 0.009043 $\uparrow$ \\ \cline{2-4}
                        &$Q_{QoS}$  &0.242338 $\pm$ 0.000516 $\uparrow$    &0.236677 $\pm$ 0.002211  \\ \hline
\multirow{3}{*}{Task5}  &$Q_{mt}$   &  0.000042 $\pm$ 0.000030     &  0.000078 $\pm$ 0.000031 $\uparrow$           \\ \cline{2-4} 
                        &$Q_{s}$    &  0.915933 $\pm$ 0.012888      &  0.928189 $\pm$ 0.002486 $\uparrow$                 \\ \cline{2-4}
                        &$Q_{QoS}$  & 0.238189 $\pm$ 0.000240$\uparrow$      &  0.237232 $\pm$  0.000391  \\ \hline                                                   
\end{tabular}
\end{table}

Second, to compare the results generated from two evaluation approaches, we demonstrate an example solution that shows the differences in composite web services obtained through two different methods. Fig. \ref{comparison} shows two composition weighted graphs as the solutions to Task 3 with $(1)$ QoS-aware approach and $(2)$ Comprehensive quality-aware method respectively. Two approaches generate exactly the same service workflow structure where those service vertices and edges denoted in red are different. We make a comparison of the quality among these different edges ($e_1$ to $e4$) associated service vertices in terms of quality of semantic matchmaking and QoS attributes. We also look at $\Delta Q$ which reveals the amount of variation on quality between two methods, where the positive values means the benefits gained under our approach while the negative values means the trade-off. To demonstrate the benefits of the positive value, we analyse the smallest positive $\Delta Q$ corresponding to $e_4$ and demonstrate how the output and required input are different under two approaches in Fig. \ref{comparisontaxonomy}. $Ser1640238160$ and $Ser1947554374$ are selected service vertices with output concept-related parameters $Inst795998200$ and $Inst582785907$ corresponding to two concepts $Con103314376$ and $Con2037585750$ respectively, and $Inst658772240$ are the required parameter related to concept $Con2113572083$. There exist $Inst795998200$ $\in$ $Con103314376$ $\sqsubseteq_{2}$ $Inst658772240$ $\in$ $Con2113572083$ and $Inst582785907$ $\in$ $Con2037585750$ $\sqsubseteq_{3}$ $Inst658772240$ $\in$ $Con2113572083$. It is obvious that our approach selects the service providing $Inst795998200$ that are closer to the users' requirements suggesting that our method can produce better semantic matchmaking quality.

@@

To analyse the effectiveness of our approach, we study the convergence rate of the proposed method in this section to understand the convergence rate of five tasks in WSC09. We analysis the performances during the whole evolutionary process in Fig. \ref{exp_fitnessvalue}, in which experiment results on five tasks are arranged in four groups consisting of average fitness, average matchType quality, average similarity quality and average QoS for generation 0-99 with optimum.

Firstly, the average fitness value with optimum is calculated by averaging the best fitness found in each generation over 30 independent runs. We can see that there is a significant increase in the fitness values towards optimum between generation 0 and generation 15-25, the remaining generation continues to produce a steady but moderate improvement in the fitness value and eventually reach a plateau with no further improvements can be observed. The same behaviour is observed over the rest tasks.

We also investigate the variation of quality of semantic matchmaking where average matchType quality with optimum and average similarity quality with optimum are studied from generation 0 to 99 in second and third column groups of Fig. \ref{exp_fitnessvalue}. In these subfigures, both the average matchType quality with optimum and average similarity quality with optimum refers to the mean values of the quality values associated to the best fitness value found in each generation over 30 independent runs. Similar to the fitness values, clear evidence of fast convergence can be observed with respect to theses quality values. This observation is also consistent for the improvement of mean QoS quality with optimum shown in the last column in Fig \ref{exp_fitnessvalue}. Additionally, we don't see too much trade-off from the QoS as the a consistently increase in semantic matchmaking quality is observed.
